MetaData262: Computing and Visualizing Test262 MetaData

### 2022-05-27 paper.tex
Last version before submition

\title{Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript}
\subtitle{}

%
\titlerunning{Live Metadata for Test262}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Diogo Costa Reis\\ist187526\\
\email{diogo.costa.reis@tecnico.ulisboa.pt}}
%
\authorrunning{Diogo Costa Reis}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Instituto Superior Técnico\\
Av. Rovisco Pais, 1\\
1049-001 Lisboa\\
Tel: +351 218 417 000\\
\email{mail@tecnico.ulisboa.pt}}
%

@@

%
\begin{abstract}
The ECMAScript standard is the specification for JavaScript which is one of the most popular programming languages in the world. Implementing the standard is challenging in various ways, both due to its complexity and its scale. Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations. There is the project MetaData262 v0 that calculates the metadata of the Test262 test suite. With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface. The web interface will also allow to filter tests by some characteristics such as built-ins and version of the standard as well as display statistics of the metadata.

@@

@@

@@

% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
JavaScript (JS) is a programming language mainly used in the development of client side web applications, also being one of the most popular programming languages. According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as second most active languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}

% Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de acordo no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
% -----------------------
% overview do standard -> descrita num standard
% Porque é muito importante que as várias implementacoes da linguagem coincidam -> o JavaScript está especificado num documento semi-formal que ...
% Falar sobre o standard
%   - o standard está como um interpretador de JavaScript em
%     pseudo-codigo - descreve detalhadamente os passos que
%    um interpretador de JS tem de executar ao avaliar qualquer
%    statement da linguagem
% Falar sobre o comité -  quem controla a evolucao do ECMAScript
ECMAScript standard\cite{ECMAScript} is the official document, written in the English language, in which the JavaScript language is defined. This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard. The standard is currently in its twelfth version. 
%
The standard specifies the \texttt{JavaScript} language, to ensure its multiple compilers and interpreters implementations are coherent. Some of the \texttt{JavaScript} compilers are the Hop~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are Node.js~\cite{Node.js} and SpiderMonkey~\cite{SpiderMonkey}. These are only four implementations among many others, which come along with the many use cases that \texttt{JavaScript} has.
%
\texttt{JavaScript} is mostly used in the web context, both client-side within browsers and server-side, but also in embedded devices. Since \texttt{JavaScript} is used in so many scenarios and across so many different contexts, it is highly important that ECMAScript standard is defined in great detail to ensure consistency. Browsers, for example, need to run \texttt{JavaScript} implementations that coincide so that websites are correctly rendered and exhibit the same behavior. In order to achieve coherent implementations, the standard defines the types, values, objects, properties, syntax, and semantics of \texttt{JavaScript} that must be the same in every \texttt{JavaScript} compiler and interpreter, while allowing \texttt{JavaScript} implementations to define additional types, values, object, properties, and functions.

@@

% MetaParagrafo tres tipos
\colorbox{orange}{The remaining} subsection provides a description of the three types of artifact described in the standard.
% QUESTION ainda nao foi lido pelo prof

@@

Figure \ref{fig:Array_pop_example} shows a simplified version of an array performing the pop function, where \texttt{(a)} and \texttt{(b)} are the before and after respectively.
Before preforming \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}. Property \texttt{length} represents the size of the array that has value \texttt{2}, while the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
After \texttt{pop} is preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}) and the \texttt{length} property (highlighted in green) is decremented by one since the size of the array changes to one.

% TODO change length with quotation marks
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.4\textwidth]{images/array_pop_example.png}
    \caption{Example Array.pop}
    \label{fig:Array_pop_example}
\end{figure}
Figure \ref{fig:Array_pop} shows a snippet of the ECMAScript standard description of the pop function in the Array Built-in. To begin with, the array will be converted to and Object using the \texttt{ToObject} function, and stored in the \texttt{O} variable (step 1).
Afterwards, the array length of the previously calculated variable will be calculated with the \texttt{LengthOfArrayLike} internal function, and storing the result in the \texttt{len} variable (step 2).
At this point there are to ways to proceed depending on the value of \texttt{len}. If the value is zero, the Array is empty, then the property \texttt{length} of \texttt{O} is set to zero and \texttt{undefined} is returned (step 3).
Otherwise, when \texttt{len} is different from zero, meaning that the Array is not empty, the Array's last element will be removed (described in Figure \ref{fig:Array_pop_example}) and returned (step 4).
To begin with, the language will assert that \texttt{len} is positive (step 4.a).
Afterwards, the \texttt{newLen} variable will store the value of \texttt{len} decremented by 1 (step 4.b).
The variable \texttt{index} will store the variable calculated in the previous step represented as a String converted with the \texttt{toString} function (step 4.c).
Then, stores the value of the \texttt{O} variable at the property corresponding to \texttt{index} in the \texttt{element} variable using the \texttt{Get} function (step 4.d).
Subsequently, deletes the previously mentioned property of the \texttt{O} variable with the \texttt{DeletePropertyOrThrow} function (step 4.e).
In addition, sets the \texttt{length} property of the \texttt{O} variable  to the \texttt{newLen} using the \texttt{Set} function (step 4.f).
Finally, returning the value of the variable \texttt{element} (step 4.g).

@@

% paragrafo - JS tem muitas particularidades que dificultam o desemvolvimento e testagem  -> É muito dificil desenvolver
% novas implemenentacoes da linguagem
% Existe uma bateria de testes que testa as implementacoes da
% linguagem contra o standard
% Esta bateria de testes é dificil de manter
% Porque? - muitos testes, muitas features, em geral ha retrocompatibilidade mas ha um pequeno numero de casos onde a retrocompatibilidade nao se verifica -> os testes de ser modificados
Implementing a \texttt{JavaScript} engine is particularly difficult since it involves dealing with the many corner cases that exist in the language. To test that corner cases are correctly dealt with there is \texttt{Test262}\cite{Test262}, the ECMAScript standard test battery. Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around \colorbox{orange}{39837} divided into \colorbox{orange}{87} subfolders, each correspond to roughly one section of the standard. \texttt{Test262} complexity grows with changes to the standard since in most cases backward compatible is maintained except for a few select cases. 
% 87 subfolders on built-ins + language
% QUESTION round number? (40000)
% QUESTION 105 subfolders if add intl402 (which also has files that were ignore)

@@

\begin{itemize}
\item \texttt{description} - contains a short description about what will be tested;
%
\item \texttt{esid} - contains the hash identifier of the ECMAScript portion associated with the feature that will be tested (the identifier references the most recent version of ECMAScript when the test is created);
%
\item \texttt{info} - contains a deeper explanation of the test behavior, frequently includes a direct citation of the standard;
%
\item \texttt{negative} - indicates that the test throws an error; associated to the keyword will be the type of error the test is supposed to be thrown (e.g. \texttt{TypeError}, \texttt{ReferenceError}) as well as the phase in which the error is expected to be thrown (e.g. \texttt{parse} vs \texttt{resolution} vs \texttt{runtime});
%
\item \texttt{includes} - contains the list of \texttt{harness} files that should be included in the execution of the tests (\texttt{Test262} makes use of a large number of auxiliary function defined in a dedicated library referred to as the \texttt{Test262} \texttt{harness} described later in this section);
%
\item \texttt{author} - contains the identification of the author of the test;
%
\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \colorbox{orange}{\texttt{Agent Record}} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
% QUESTION clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
%
%
\item \texttt{features} - contains a list of features that are used in the test;
%
\item \texttt{es5id} and \texttt{es6id} - indicates that the feature being tested belongs to ECMAScript 5 and 6 respectively and contains the hash identifier of the section of the standard it belongs to; these keywords have been deprecated and substituted by esid.
\end{itemize}

% Varias observacoes: 
%  - A metada esta muitas vezes INCOMPLETA e algumas vezes incorrecta 
%  - é preciso calcular a metadata certa
As Figure 5 illustrates, it is often the case that the metadata of a test is incomplete. Some tests also have the wrong metadata. As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata as well as completing the metadata that is \colorbox{orange}{missing.}

@@

%% metadados que achamos relevantes e nao estao incluidos
%% ...
\colorbox{orange}{The example} in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id. Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be useful to have \texttt{includes} with the \texttt{harness} files needed to execute the test. The \texttt{harness} information is very useful since it makes it easy to identify the part of the \texttt{harness} needed to run that test, opening the door for loading only part of the \texttt{harness} instead of the whole \texttt{harness} which is the current approach.

@@

Scientific studies about analysis of programs and techniques of instrumentation of JavaScript encompasses a varied range of specifications such as: type systems (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), abstract interpreters (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), \emph{points-to} analysis (\cite{Points-to_analysis_for_JavaScript}), Hoare logic (\cite{Javert-2017}, \cite{Javert-2019}), operational semantics (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), intermediate representations and compilers (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}). Next, we have the related work  most relevant for this thesis, it is organized into two Tables \ref{tab:related-work_timeline-1} and \ref{tab:related-work_timeline-2}. This time line focuses on the most important investigation work, focused on the JavaScript language specification since its creation in 1995.

% https://tex.stackexchange.com/a/196808
\begin{table}[h!]
\vspace{-2em}
\renewcommand{\arraystretch}{1.8}

\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
  1995 & JavaScript is launched and implemented in Netscape Navigator 2.0 beta 3.\\
  2005 & Thiemann presents the first try at defining a system of types possible for JavaScript ~\cite{Thiemann-2005}. The system detects the possible characteristics of an object and signals suspected type conversions. The work done does not cover important aspects of JavaScript, such as inheritance based on prototypes, and was not created along with an implementation at the time of the publication.\\
  2008 & Maffeis et al. proposes a first operational semantics for the language ECMAScript version 3. (ES3)~\cite{operational_semantics_javascript-2008}. This first semantics was used to analyze various properties of security of web applications.\\
  2010 & Guha et al.~\cite{essence_javascript-2010} defines a language $\lambda$JS, a formal calculation that encapsulates the fundamental characteristics of ES3. $\lambda$JS comes along an interpreter \emph{Racket}~\cite{Racket} and a translator from ES3 programs to expressions $\lambda$JS.\\
  2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
\end{tabular}
\vspace{1em}
\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
\label{tab:related-work_timeline-1}
\end{table}
\vspace{-2em}

\begin{table}[ht!]
\renewcommand{\arraystretch}{1.8}
\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}

>>>>

\title{MetaData262: Computing and Visualizing Test262 MetaData}
\subtitle{}

%
\titlerunning{MetaData262: Computing and Visualizing Test262 MetaData}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Diogo Costa Reis\\ist187526\\
\email{diogo.costa.reis@tecnico.ulisboa.pt}}
%
\authorrunning{Diogo Costa Reis}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Instituto Superior Técnico\\
Av. Rovisco Pais, 1\\
1049-001 Lisboa\\
Tel: +351 218 417 000\\
\email{mail@tecnico.ulisboa.pt}}
%

@@

%
\begin{abstract}
The ECMAScript standard is the specification of the Java-Script language which is one of the most popular programming languages in the world. Implementing the ECMAScript standard is challenging in various ways, both due to its complexity and its scale. Test262 was created to alleviate this difficulty by providing a comprehensive test suite that allows developers to test their implementations. 
%
However, Test262 does not come to a structured description of its tests, making it impossible to filter tests by the JavaScript features they target. The MetaData262 v0 system was created to help with this issue, providing a database containing the metadata of Test262 tests and a collection of scripts to compute that metadata.
%
With this thesis, we will improve MetaData262 v0 in two separate ways. First, we will re-engineer the system for metadata computation, making it more precise and performant and enriching the computed metadata with additional information. Second, we will develop a web interface for the visualization of Test262, allowing developers to easily filter tests by feature.

@@

% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
JavaScript (JS) is a programming language mainly used in the development of client-side web applications, also being one of the most popular programming languages. According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as the second most active language on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}

% Para 2 --> implemetar JS e' dificil: complexidade tamanho
% Para 2.5 --> Test262 ajuda a implementar permitindo testar implementacoes --> test262 tambem e' grande dificil selecionar os testes apropriados
JavaScript is defined in the \texttt{ECMAScript} standard~\cite{ECMAScript}, a document written in English. The standard has been growing with each new version, both in size and complexity. The growth of the standard is illustrated in Figure \ref{fig:paginas_standard_version}. This continuous growth in complexity makes the JavaScript a language very hard to implement with many corner cases. \texttt{Test262}~\cite{Test262} was created to aid the development of JavaScript engines by providing a conformance test suite that developers can use to test their implementations. As the standard is extensive and complex, a conformance test suite for JavaScript must necessarily be very large given that it must cover all the corner cases of the standard. The standard being complex and extensive also leads to many of its implementations being only partial, particularly those made in academic contexts, for example JSRef~\cite{JSCert-2014} and JISET~\cite{JISET-2020}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.7\textwidth]{images/standard_pages_evolution.png}
    \caption{Evolution of ECMAScript standard's number of pages}
    \label{fig:paginas_standard_version}
\end{figure}

% Para 4 --> MetaData v0 calcula metadata --> disponibiliza mongoDB

When implementing a JavaScript engine or a JavaScript analysis, it is very important to know the tests that should pass for the specific feature being implemented. The selection of applicable tests for partial implementations is particularly difficult since tests for one feature may or may not use other features that are not implemented. For instance, a test targeting the Math library might use features of the array library. \texttt{MetaData262 v0} is a project developed at IST with the goal of facilitating this problem. In particular, MetaData262 v0 processes the \texttt{Test262} test suite and calculates metadata about the tests, storing that metadata in a NoSQL database that can be easily inspected. \texttt{MetaData262 v0} stores for each test various properties, such as the JavaScript version to which the test belongs, the syntactic constructs that are used in the test, and the built-ins that are used by the test.

% --> eu vou melhorar metadata
% Para 5 --> Visualization system permite interagir com a metadata via web com interface simples --> filtrar; verdados varias maneiras; estatisticas
In this thesis we will improve \texttt{MetaData262 v0} in two ways. First, we will develop a \texttt{MetaData262 v1}, a new version of MetaData262 with an improved infrastructure for computing the metadata of Test262 tests, making it more precise and extending it with additional information. Second, we will build \texttt{MetaData262Viz}, a web application for the visualization of the metadata that will allow the developers to easily select tests per feature and obtain statistics regarding the selected testes.

% Para 5.1 --> improve MetaData262 v0
As described above, MetaData262 v1 will improve MetaData262 v0 in the following ways. Below, we give a brief overview of the most relevant improvements:

\begin{itemize}
\item \emph{Better Precision:} MetaData262 v0 computes for each test the version to which the test belongs and the built-in objects with which the test interacts. Unfortunately, this computation is far from precise as it relies on the specific implementations of JavaScript engines. For instance, MetaData262 v0 determines that a test is meant for version 5 if a JS engine implementing version 5 passes the test. Naturally, this approach is not error-free since it is never the case that a JS engine implements one and only one version of the language fully. To improve precision, we will build a dedicated infrastructure to streamline the addition of new JavaScript engines to our metadata computation pipeline and use this new infrastructure to compute a new version of the metadata.
%
\item \emph{Better Performance} As said before, the metadata computation pipeline of MetaData262 v0 requires running various JS engines on each test of Test262. This is extremely time consuming given the high number of tests contained in Test262 (around 40K). 
Currently this is done sequentially. To improve on this situation, we will associate each JavaScript engine with a docker container and run the docker containers in parallel. With docker containers running in parallel the scalability and performance of \texttt{MetaData262 v1} will improve. In addiction to that, the JavaScript engines being associated with docker containers will also make \texttt{MetaData262 v1} easier to replicate.
%
\item \emph{More information:} We will extend the computed metadata of each test with additional information. In particular, we plan to associate with each test: (1) the date in  which the test was introduced and (2) the harness files and functions that the test uses. Basically, the Test262 harness is a set of files that contain auxiliary functions used by Test262 tests, such as assert and assertEquals. 
\end{itemize}

% Para 5.2 --> build MetaData262 visualization system
MetaData262Viz will be a visualization system for the metadata computed with MetaData262 v1. In particular, MetaData262Viz will allow developers to filter Test262 tests by feature and to obtain statistics about the selected tests. The statistics will generate charts built in real time with the metadata of the tests selected. Some examples of those charts to generate are a bar charts with the amount of tests per version and a pie chart with the the amount of tests per built-in used.
We have already developed a prototype of MetaData262Viz built with Node.js~\cite{Node.js} and React~\cite{React}. The prototype already allows filtering the tests by their version and the built-ins used. As for the output, it can already display the full metadata or just the path of the tests selected.

% Para 6 --> resumo do resto do relatorio
This document is structured in the following way. Section \ref{sec:Goals} has the objectives of this thesis. Section \ref{sec:Background} gives an overview of the field of the thesis, describing the \texttt{ECMAScript} standard, the \texttt{Test262} test suite, and the project \texttt{MetaData262 v0}. Next, Section \ref{sec:Related Work} presents the scientific work done that relates to this thesis topic. Section \ref{sec:Design and Methodology} presents the design and methodology that will be be followed in this thesis. As for Section \ref{sec:Evaluation and Planning}, it describes how we plan to evaluate the results of this thesis as well as its timeline. Finally, Section \ref{sec:Conclusion} presents the main conclusions and a summary of this proposal.

@@

This thesis has two main goals. The first is to improve \texttt{MetaData262 v0} and the second is to create a web visualization system to make the metadata more accessible.
%
In the end, we hope that developers of JS engines and analyses will use the MetaData262Viz for filtering Test262 tests per feature, making the test selection process more efficient, effective, and reliable.

\paragraph{MetaData262 v0 improvements}
This goal is to improve the precision and performance of MetaData v0 making it more complete, scalable, and easier to replicate.

\paragraph{MetaData262Viz}
This goal aims to create a web platform to easily visualize the metadata calculated by \texttt{MetaData262 v0}. The system will provide filters with which the developer will be able to filter Test262 tests by their features. Then, the system will output the filtered tests in various formats as well as useful statistics about them.

@@

% Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de acordo no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
% -----------------------
% overview do standard -> descrita num standard
% Porque é muito importante que as várias implementacoes da linguagem coincidam -> o JavaScript está especificado num documento semi-formal que ...
% Falar sobre o standard
%   - o standard está como um interpretador de JavaScript em
%     pseudo-codigo - descreve detalhadamente os passos que
%    um interpretador de JS tem de executar ao avaliar qualquer
%    statement da linguagem
% Falar sobre o comité -  quem controla a evolucao do ECMAScript
ECMAScript standard is the official document, written in the English language, in which the JavaScript language is defined. This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard. The standard is currently in its twelfth version. 
%
The standard specifies the \texttt{JavaScript} language, to ensure its multiple compilers and interpreters implementations are coherent. Some of the \texttt{JavaScript} compilers are the Hop~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are Node.js and SpiderMonkey~\cite{SpiderMonkey}. These are only four implementations among many others, which come along with the many use cases that \texttt{JavaScript} has.
%
\texttt{JavaScript} is mostly used in the web context, both client-side within browsers and server-side, but also in embedded devices. Since \texttt{JavaScript} is used in so many scenarios and across so many different contexts, it is highly important that ECMAScript standard is defined in great detail to ensure consistency. Browsers, for example, need to run \texttt{JavaScript} implementations that coincide so that websites are correctly rendered and exhibit the same behavior. In order to achieve coherent implementations, the standard defines the types, values, objects, properties, syntax, and semantics of \texttt{JavaScript} that must be the same in every \texttt{JavaScript} compiler and interpreter, while allowing \texttt{JavaScript} implementations to define additional types, values, object, properties, and functions.

@@

% MetaParagrafo tres tipos
The remaining subsection provides a description of the three types of artifact described in the standard.

@@

Figure \ref{fig:Array_pop_example} shows a simplified version of an array performing the pop function, where \texttt{(a)} and \texttt{(b)} are the before and after respectively.
Before performing \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}. Property \texttt{length} represents the size of the array that has value \texttt{2}, while the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
After \texttt{pop} is performed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}) and the \texttt{length} property (highlighted in green) is decremented by one since the size of the array changes to one.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.4\textwidth]{images/array_pop_example.png}
    \caption{Example Array.pop}
    \label{fig:Array_pop_example}
\end{figure}
Figure \ref{fig:Array_pop} shows a snippet of the ECMAScript standard description of the pop function in the Array Built-in. To begin with, the array will be converted to and Object using the \texttt{ToObject} function, and stored in the \texttt{O} variable (step 1).
Afterwards, the array length of the previously calculated variable will be calculated with the \texttt{LengthOfArrayLike} internal function, and storing the result in the \texttt{len} variable (step 2).
At this point there are to ways to proceed depending on the value of \texttt{len}. If the value is zero, the Array is empty, then the property \texttt{length} of \texttt{O} is set to zero and \texttt{undefined} is returned (step 3).
Otherwise, when \texttt{len} is different from zero, meaning that the Array is not empty, the Array's last element will be removed (described in Figure \ref{fig:Array_pop_example}) and returned (step 4).
To begin with, the language will assert that \texttt{len} is positive (step 4.a).
Afterwards, the \texttt{newLen} variable will store the value of \texttt{len} decremented by 1 (step 4.b).
The variable \texttt{index} will store the variable calculated in the previous step represented as a String converted with the \texttt{toString} function (step 4.c).
Then, stores the value of the \texttt{O} variable at the property corresponding to \texttt{index} in the \texttt{element} variable using the \texttt{Get} function (step 4.d).
Subsequently, deletes the previously mentioned property of the \texttt{O} variable with the \texttt{DeletePropertyOrThrow} function (step 4.e).
In addition, sets the \texttt{length} property of the \texttt{O} variable  to the \texttt{newLen} using the \texttt{Set} function (step 4.f).
Finally, returning the value of the variable \texttt{element} (step 4.g).

@@

% paragrafo - JS tem muitas particularidades que dificultam o desemvolvimento e testagem  -> É muito dificil desenvolver
% novas implemenentacoes da linguagem
% Existe uma bateria de testes que testa as implementacoes da
% linguagem contra o standard
% Esta bateria de testes é dificil de manter
% Porque? - muitos testes, muitas features, em geral ha retrocompatibilidade mas ha um pequeno numero de casos onde a retrocompatibilidade nao se verifica -> os testes de ser modificados
Implementing a \texttt{JavaScript} engine is particularly difficult since it involves dealing with the many corner cases that exist in the language. To test that corner cases are correctly dealt with there is \texttt{Test262}, the ECMAScript standard test battery. Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around 40000 divided into 87 subfolders, each correspond to roughly one section of the standard. \texttt{Test262} complexity grows with changes to the standard since in most cases backward compatible is maintained except for a few select cases. 
% 87 subfolders on built-ins + language
% QUESTION round number? (40000)
% QUESTION 105 subfolders if add intl402 (which also has files that were ignore)

@@

\begin{itemize}
\item \texttt{description} - contains a short description about what will be tested;
%
\item \texttt{esid} - contains the hash identifier of the ECMAScript portion associated with the feature that will be tested (the identifier references the most recent version of ECMAScript when the test is created);
%
\item \texttt{info} - contains a deeper explanation of the test behavior, frequently includes a direct citation of the standard;
%
\item \texttt{negative} - indicates that the test throws an error; associated to the keyword will be the type of error the test is supposed to be thrown (e.g. \texttt{TypeError}, \texttt{ReferenceError}) as well as the phase in which the error is expected to be thrown (e.g. \texttt{parse} vs \texttt{resolution} vs \texttt{runtime});
%
\item \texttt{includes} - contains the list of \texttt{harness} files that should be included in the execution of the tests (\texttt{Test262} makes use of a large number of auxiliary function defined in a dedicated library referred to as the \texttt{Test262} \texttt{harness} described later in this section);
%
\item \texttt{author} - contains the identification of the author of the test;
%
\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \texttt{Agent Record} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
% QUESTION clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
%
%
\item \texttt{features} - contains a list of features that are used in the test;
%
\item \texttt{es5id} and \texttt{es6id} - indicates that the feature being tested belongs to ECMAScript 5 and 6 respectively and contains the hash identifier of the section of the standard it belongs to; these keywords have been deprecated and substituted by esid.
\end{itemize}

% Varias observacoes: 
%  - A metada esta muitas vezes INCOMPLETA e algumas vezes incorrecta 
%  - é preciso calcular a metadata certa
As Figure 5 illustrates, it is often the case that the metadata of a test is incomplete. Some tests also have the wrong metadata. As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata as well as completing the metadata that is missing.

@@

%% metadados que achamos relevantes e nao estao incluidos
%% ...
The example in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id. Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be useful to have \texttt{includes} with the \texttt{harness} files needed to execute the test. The \texttt{harness} information is very useful since it makes it easy to identify the part of the \texttt{harness} needed to run that test, opening the door for loading only part of the \texttt{harness} instead of the whole \texttt{harness} which is the current approach.

@@

There is a wide range of scientific studies on program analysis for JavaScript, including: type systems (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), abstract interpreters (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), \emph{points-to} analysis (\cite{Points-to_analysis_for_JavaScript}), Hoare logic (\cite{Javert-2017}, \cite{Javert-2019}), operational semantics (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), intermediate representations and compilers (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}). Here, we focus only on the related work most relevant for this thesis, covering all the most relevant academic efforts in modelling the JavaScript semantics.

One of the first formal semantics of the JS language was presented by Thiemann~\cite{Thiemann-2005} as part of his work on defining a type system for JavaScript. Thiemann's semantics targets the third version of the language and is specified in small-step style. Importantly, this semantics ignores a number of important features of the JavaScript language, such as implicit type coercion, the treatment of global variables, and the non-strict mode.

In 2008, Maffeis et al. proposed the first operational semantics for a significant subset of ECMAScript~\cite{operational_semantics_javascript-2008}. They target the 3rd version of the language. This semantics was formalized in big-step style and comprises a large collection of rules. Unfortunately, the semantics has not been automated, meaning that it cannot be tested. Nevertheless, the authors use the proposed semantics to reason about various security properties of JS web applications.

In 2010, Guha et al define lambdaJS~\cite{essence_javascript-2010}, a lambda calculus to be used as a compilation target for JavaScript. The authors claim that lambdaJS capture the fundamental behavior of the JS language. The authors also define a desugaring function that compiles ES 3 programs to lambdaJS and implement a lambdaJS interpreter in Racket. However, the authors do not test their infrastructure against Test262. In order to test this work against Test262, one would have to compile all Test262 tests to lambdaJS and then run the compiled test suite using the lambdaJS interpreter. Importantly, the lambdaJS infrastructure does not support the entire ES3 standard. Hence, if one was to test the proposed infrastructure, one would first have to clearly define the target subset of ES3 and filter the appropriate tests from Test262. Later, the same authors propose S5~\cite{Politz-2012}, an extension of lambdaJS to the 5th version of the standard. In contrast to lambdaJS, which the authors did not test, the authors do test S5, running it against Test262. They pass a total of 8157 tests out of a total of 11606 tests (approx. 70\%). However, the authors admit that they do not implement the whole language, stating that they only implement 60\% of the built-in objects. Nevertheless, they were not able to filter the tests appropriately to include only the tests with the features that they do implement. This means that they might have bugs in features that they implement that were disregarded as being caused by unimplemented features. Without an appropriate labeling of the tests, the only way of guaranteeing that they are not failing tests that they should be passing is to analyze all failing tests manually, which the authors do not do.
